### 2024-11-08 docs/paper.tex
Removed a ton of double spaces
A ew minor tweaks

\maketitle
\begin{abstract}
{\begin{spacing}{1.2}
Housing is the largest asset in U.S. household portfolios, and a first-time homebuyers increasingly rely on parental transfers. This paper quantifies the contribution of parental transfers to the homeownership rate of young households. I build and estimate a life-cycle overlapping generations model with housing, where adult children and parents interact without commitment. I find that parental transfers account for 14 percentage points (29\%) of young households' homeownership. Transfers from wealthy parents not only help households overcome borrowing constraints but also help them maintain homeownership, reducing the downsides of illiquidity. Policies lowering entry barriers to homeownership generally increase the reliance on parental wealth, while increasing liquidity lowers the role of parental wealth. Finally, I study how illiquidity affects the commitment problem and find that children with wealthy parents use the illiquidity of housing to encourage transfers and exhibit a preference for illiquidity.
\end{spacing}}
\end{abstract}
\newpage
\onehalfspacing

@@

Next, I study how housing illiquidity affects the strategic behavior arising in the absence of commitment. I find that 27\% of young households prefer that housing is illiquid---due to sales costs---rather than liquid. These households have wealthy parents and use the sales costs as a commitment device to maintain low liquid wealth in the next period, effectively becoming ``wealthy hand-to-mouth''. Lower liquid wealth implies higher marginal propensities to consume (MPC), and hence more parental transfers. As an illustrative example, consider a child who buys a house and brings no liquid wealth to the next period. If the parent does not transfer in the next period, the child will either face very low consumption (with a correspondingly high marginal value of wealth) or be forced to liquidate the house and incur sales costs. The parent dislikes both outcomes, and may find it optimal to transfer enough to keep the child in the home. This driver of preference for illiquidity differs from those studied in the literature, such as temptation or time inconsistency \citep{attanasio2024temptation,laibson1997golden}. Furthermore, that parental wealth generates preference for illiquidity provides a novel explanation for the existence of high-MPC households \cite[see e.g.,][]{kaplan2022marginal}.

@@

My paper also contributes to the literature on altruistic households who interact without commitment by studying how illiquid housing affects the commitment problem. In these models children have an incentive to undersave to increase parental transfers (e.g., \citealp{Altonji1997a,Boar2018,Barczyk2014,Chu2020}). However, the illiquidity of housing imposes future expenditure commitments \citep{Chetty2007,Shore2010}. The illiquidity of housing allows households to commit to having low liquid wealth, which increases future transfers. The paper most closely related to mine is \cite{Barczyk2020a}, which studies the mirror problem: how the homeownership of retired parents affects the economic behavior of the elderly and their adult children. In their model, the parent’s home serves as a commitment device, encouraging wealth bequests from the parent and informal care from the adult offspring.

>>>>

\maketitle
\begin{abstract}
{\begin{spacing}{1.2}
Housing is the largest asset in U.S. household portfolios, and first-time homebuyers increasingly rely on parental transfers. This paper quantifies the contribution of parental transfers to the homeownership rate of young households. I build and estimate a life-cycle overlapping generations model with housing, where adult children and parents interact without commitment. I find that parental transfers account for 14 percentage points (29\%) of young households' homeownership. Transfers from wealthy parents not only help households overcome borrowing constraints but also help sustain homeownership, reducing the downsides of illiquidity. Policies lowering entry barriers to homeownership generally increase the reliance on parental wealth, while increasing liquidity lowers the role of parental wealth. Finally, I show that children of wealthy parents strategically use the illiquidity of housing as a commitment device to encourage transfers and exhibit a preference for illiquidity.
\end{spacing}}
\end{abstract}
\newpage
\onehalfspacing

@@

I then study how housing illiquidity affects the strategic behavior arising in the absence of commitment. I find that 27\% of young households prefer illiquid housing---due to sales costs---to liquid housing. These households have wealthy parents and use the sales costs as a commitment device to maintain low liquid wealth in the next period, effectively becoming ``wealthy hand-to-mouth.'' Lower liquid wealth implies higher marginal propensities to consume (MPC), and hence more parental transfers. As an illustrative example, consider a child who buys a house and brings no liquid wealth to the next period. If the parent does not transfer in the next period, the child will either face very low consumption (with a correspondingly high marginal value of wealth) or be forced to liquidate the house and incur sales costs. The parent dislikes both outcomes, and may find it optimal to transfer enough to keep the child in the home. This driver of preference for illiquidity differs from those studied in the literature, such as temptation or time inconsistency \citep{attanasio2024temptation,laibson1997golden}. Furthermore, that parental wealth generates preference for illiquidity provides a novel explanation for the existence of high-MPC households \cite[see e.g.,][]{kaplan2022marginal}.

@@

My paper also contributes to the literature on altruistic households interacting without commitment by studying how illiquid housing affects the commitment problem. In these models children have an incentive to undersave to increase parental transfers (e.g., \citealp{Altonji1997a,Boar2018,Barczyk2014,Chu2020}). Illiquid housing imposes future expenditure commitments \citep{Chetty2007,Shore2010}, which children of wealthy parents strategically exploit to encourage future parental transfers. The paper most closely related to mine is \cite{Barczyk2020a}, which studies the mirror problem: how the homeownership of retired parents affects the economic behavior of the elderly and their adult children. In their model, the parents' home serves as a commitment device, encouraging wealth bequests from the parent and informal care from the adult offspring.
